Salt rejection through Electrostatically Gated Nanoporous Graphene

### 2019-05-16 paper/SI.tex
Add converting script and SI
Add SI before the correction

\documentclass[manuscript=suppinfo,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\newcommand{\Fig}{Fig.}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\theequation}{S\arabic{equation}}
\SectionNumbersOn
\usepackage{xr-hyper}
\usepackage{xcolor}
\externaldocument[main-]{paper}

@@

\author{Hyung Gyu Park} \affiliation{Nanoscience for Energy Technology and Sustainability,
Department of Mechanical and Process Engineering,
Eidgenössische Technische Hochschule (ETH) Zürich,
Tannenstrasse 3, Zürich CH-8092, Switzerland.}
\alsoaffiliation{Mechanical Engineering, Pohang University of Science and
  Technology (POSTECH), 77 Cheongam-ro, Nam-gu, Pohang, Gyeongbuk,
  Republic of Korea.}
\email{iduserpark@gmail.com}

@@

\title{Macroscopic Salt Rejection through Electrostatically Gated Nanoporous Graphene}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    Fe{[(CN)$_{6}$]}$^{3-}$   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical
    diffusion through PCTE can be calculated using Eq. 1 in the main
    text and the diffusivities of the respective anion and cation.}
  \label{tab:diff}
\end{table}

The conductivity increase $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity \textit{versus} concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$L) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 1.51$\pm$0.27$\times$10$^{-2}$  & 106   & 1.34 $\pm$ 0.24\\
    NaCl        & 9.90$\pm$0.17$\times$10$^{-3}$ & 91    & 1.02 $\pm$ 0.02\\
    LiCl        & 6.03 $\pm$0.80$\times$10$^{-3}$        & 83    & 0.682 $\pm$ 0.090\\
    CaCl$_{2}$  & 1.25$\pm$0.1$\times$10$^{-2}$    & 149   & 0.789 $\pm$ 0.063\\
    MgSO$_{4}$  & 8.80$\pm$1.2$\times$10$^{-3}$   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 2.40$\pm$0.28$\times$10$^{-2}$   & 395   & 0.575 $\pm$ 0.063\\
    K$_{2}$SO4$_{4}$    & 1.98$\pm$0.20$\times$10$^{-2}$  & 258   & 0.722 $\pm$ 0.073\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{img/SI-eta-all.pdf}
  \caption{Measured $\eta$ values as function of $V_{\mathrm{G}}$ for
    different salts used in this work. The concentration in HCR is 0.1
    mM in all cases.}
  \label{fig:eta-all}
\end{figure}

\section{Control experiment with copper tape}
\label{sec:copper}
We make sure that the applied voltage is effectively applied via the
graphene membrane and not through leakage directly coupling the
copper tape to the ionic solution. For this, we use a device as in \Fig
1a, omitting the graphene and measuring the current through the
membrane with the Autolab electrochemical workstation using the
membrane as working electrode and a platinum wire as counter/reference
electrode 3.5 cm apart in 0.1 mM KCl. When no graphene is inserted in
the device, a current in the baseline range ($\sim$ 0.1 nA) is
observed, indicating no current passing from the membrane. However, if
graphene is inserted, a current in the range of $\sim$ 10 - 100 nA is
measured. If the copper tape is in contact with the solution when
omitting the Kapton tape, the current increases to $\sim$ 1 - 10
$\mathrm{\mu}$A, independent of the exposed area of the copper tape to
the solution.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-resistance-rectification.pdf}
  \caption{Simple circuit analog model for the transport through
    PG-PCTE membrane. \textbf{a}. Schematic of PG-PCTE membrane in
    cross-section view. \textbf{b}. Model for transport resistance in
    analogy to an electric circuit. This schematic depicts a cut
    through a PG-PCTE membrane showing a single PCTE pore and multiple
    graphene pores.}
  \label{fig:R-model}
\end{figure}
The ionic transport through the PG-PCTE membrane is the concerted
effect the transport resistance through PG and PCTE membranes,
respectively. As schematically shown in \Fig{} \ref{fig:R-model}, the
overall transport resistance of the system is composed of the series
resistance $R_{\mathrm{G}}$ and $R_{\mathrm{PCTE}}$, respectively. For
both PG and PCTE membranes, the resistance $R$ is defined as the ratio
between the concentration between HCR and LCR $\Delta c$, and the total diffusive current $I$ as:
\begin{equation}
\label{eq:1}
R_{i} = \Bigg|\frac{\Delta c}{I_{i}}\Bigg| =  \Bigg|\frac{\Delta c}{J_{i}A_{i}}\Bigg|
\end{equation}
where $i$ denotes the membrane type and $A_{i}$ is the area of the
pore. We further treat $R_{\mathrm{G}}$ as the effective resistance of
$\gamma$ parallel resistors with individual resistance
$R_{\mathrm{g}}$, where $\gamma$ is the average number of PG pores per
PCTE channel, and $R_{\mathrm{g}}$ is the average transport resistance
per PG pore. The value of $\gamma$ is determined from geometric
analysis of both membranes: given the pore number density of PG
($\sim$1.25$\times$10$^{14}$ m$^{-2}$), the pore number density of
PCTE ($\sim$1.50$\times$10$^{12}$ m$^{-2}$), and the percentage of
PCTE pores with PG on top ($\sim$0.13), $\gamma$ is determined to be
$\sim$10.83.

The objective of the circuit analog model is to find effective salt
rejection ratio through PG, $\xi$ extracted from the experimental
$\eta$ values (defined in main text Eq. \ref{main-eq:rejection}). By
defining $R_{\mathrm{G}}^{0}=R_{\mathrm{G}}(V_{\mathrm{G}}=0)$,
$\delta = R_{\mathrm{PCTE}} / R_{\mathrm{G}}^{0}$,
$\chi = R_{\mathrm{G}}/R_{\mathrm{G}}^{0}$, $\eta$ is rewritten using
transport resistances as:
\begin{equation}
  \begin{aligned}[t]
      \label{eq:eta-R}
  \eta &= 1 - {\displaystyle
    \frac{(R_{\mathrm{PCTE}} + R_{\mathrm{G}})^{-1}}
    {(R_{\mathrm{PCTE}} + R_{\mathrm{G}}^{0})}} \\
    &= 1 - \frac{\delta + 1}{\delta + R_{\mathrm{G}}/R_{\mathrm{G}}^{0}} \\
    &= 1 - \frac{\delta + 1}{\delta + \chi}
  \end{aligned}
\end{equation}
Consequentially, the
salt rejection ratio $\xi$, is associated with $\chi$ by
$\xi = 1 - \chi^{-1}$ and is related with $\delta$ and $\eta$ via:
\begin{equation}
  \label{eq:xi-deriv}
  \begin{aligned}[t]
    \xi &= 1 - \chi^{-1} \\
    &= \frac{(\delta + 1) \eta}{\delta\eta + 1}
  \end{aligned}
\end{equation}
which is main text Eq. \ref{main-eq:xi-def}. When $0<\eta<1$ (salt
rejection regime) and $\delta>0$, Eq. \ref{eq:xi-deriv} always
guarantees $\xi>\eta$, since $\delta + 1$ is always larger than
$\delta\eta + 1$. In other words, due to the existence of
$R_{\mathrm{PCTE}}$, the effective salt rejection through PG upon
gating, is always larger than that experimentally observed.

Next we seek the value of $\delta$, the resistance ratio
between the PCTE and PG membranes, without gating. The range
of $\delta$ can be estimated from experimental data in main text
\Fig{} \ref{main-fig:2}b using the relation:
\begin{equation}
  \label{eq:exp-delta}
  \delta = \left(\frac{\boldsymbol{J}_{\mathrm{PCTE}}}{\boldsymbol{J}_{\mathrm{PG}}^{0}} -1 \right)^{-1}
\end{equation}
by assuming the validity of the circuit analog model for all salt
systems. As shown in \Fig{} \ref{fig:delta-compare}, the
experimentally estimated value of $\delta$ varies from 1 to 5,
indicating that the resistance of the PG membrane is dominating in the
PG-PCTE system.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.55\linewidth]{img/SI-delta-estimate.pdf}
  \caption{Comparison between the $\delta$ value estimated from simple
    geometric model (Eq. \ref{eq:delta-resistance}) and from
    experimental value (Eq. \ref{eq:exp-delta}), showing good
    agreement between the two methods.  }
  \label{fig:delta-compare}
\end{figure}
$\delta$ can also be estimated using geometric parameters of both
membranes. For a long PCTE channel with length $L$ and
$r_{\mathrm{PCTE}}$, $R_{\mathrm{PCTE}}$ has power law of
$L/r_{\mathrm{PCTE}}^{2}$, while for a graphene pore
with radius $r_{\mathrm{G}}$ and vanishing thickness, the resistance of individual pore $R_{\mathrm{g}}^{0}$
has power law $1/r_{\mathrm{G}}$ \cite{O_Hern_2014}, which are expressed as:
\begin{subequations}
\begin{eqnarray}
  \label{eq:R-both}
  R_{\mathrm{PCTE}} &= &k_{\mathrm{PCTE}} \frac{L}{r_{\mathrm{PCTE}}^{2}} \\
  R_{\mathrm{g}}^{0} &= &k_{\mathrm{g}} \frac{1}{r_{\mathrm{G}}}
\end{eqnarray}
\end{subequations}
where $k_{\mathrm{PCTE}}$ and $k_{\mathrm{g}}$ are the coefficient
associated with PCTE and PG membranes, respectively. The value of
$\delta$ is then expressed as:
\begin{equation}
  \label{eq:delta-resistance}
  \delta
  = \frac{R_{\mathrm{PCTE}}}{R_{\mathrm{G}}^{0}}
  = {\displaystyle
    \frac{k_{\mathrm{PCTE}}}{k_{\mathrm{g}}}
    \frac{L r_{\mathrm{G}} \gamma}
        {r_{\mathrm{PCTE}}^{2}}}
\end{equation}
Here we take $L$=24 $\mu$m, $r_{\mathrm{G}}$=10 nm,
$r_{\mathrm{PCTE}}$=200 nm. From simple Fick's law\cite{Beers_2006},
the coefficient $k_{\mathrm{PCTE}}$ is determined as
$1/(D_{\mathrm{\pm}} \pi)$. On the other hand $k_{\mathrm{G}}$ varies
by the choice of model, ranging from $3 \pi / D_{\mathrm{\pm}}$ using
the analog of the Sampson formula for
fluid\cite{Roscoe_1949,Celebi_2014}, to $1/(2 D_{\mathrm{\pm}})$ from
the Hill formula for disk absorption model
\cite{Hill_1975,Grebenkov_2018}. As a result the theoretical value of
$\delta$ has lower bound of $\sim$2.2 (Sampson formula, orange line in
\Fig{} \ref{fig:delta-compare}) to $\sim$40.5 (Hill formula). We note
our experimental values of $\delta$ is in good agreement estimated
from the Sampson formula (which we use in the main text) while lower
than the Hill formula. Having a higher estimated value of $\delta$
could yield unreasonably high value of $\xi$ and the influence of
concentration is deviated from our theoretical predictions. Further
theoretical analysis taking nonidealities including surface adsorption
and chemical nature of pore edge, is required to have better
understanding of graphene's transport resistance.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    $c_{0}$ (mM) & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm)\\
    \hline
    0.1&        1.90$\times$10$^{-2}$ &        30.4\\
    0.33&       5.41$\times$10$^{-2}$ & 16.0\\
    1   &1.33$\times$10$^{-1}$ & 10.0\\
    3.3&     5.13$\times$10$^{-1}$ &  6.0\\
    10&      1.08    &3.3\\
    \hline
  \end{tabular}
  \caption{Conductivity and calculated Debye lengths of various KCl
    concentrations in order to obtain rectification versus
    concentration curves. }
  \label{tab:debye}
\end{table}

@@

>>>>

\documentclass[manuscript=suppinfo,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\newcommand{\Fig}{Fig.}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\theequation}{S\arabic{equation}}
\SectionNumbersOn
\usepackage{xr-hyper}
\usepackage{xcolor}
\usepackage{units}
\externaldocument[main-]{paper}

@@

\author{Hyung Gyu Park} \affiliation{Nanoscience for Energy Technology and Sustainability,
Department of Mechanical and Process Engineering,
Eidgenössische Technische Hochschule (ETH) Zürich,
Tannenstrasse 3, Zürich CH-8092, Switzerland.}
\alsoaffiliation{Mechanical Engineering, Pohang University of Science and
  Technology (POSTECH), 77 Cheongam-ro, Nam-gu, Pohang, Gyeongbuk, 37673,
  Republic of Korea.}
\email{iduserpark@gmail.com}

@@

\title{Salt rejection through Electrostatically Gated Nanoporous Graphene}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{-9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    Fe{[(CN)$_{6}$]}$^{3-}$   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivity values for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical
    diffusion through PCTE can be calculated using Eq. 1 in the main
    text and the diffusivities of the respective anion and cation.}
  \label{tab:diff}
\end{table}

The conductivity increase $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity \textit{versus} concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$L) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 1.51$\pm$0.27$\times$10$^{-2}$  & 106   & 1.34 $\pm$ 0.24\\
    NaCl        & 9.90$\pm$0.17$\times$10$^{-3}$ & 91    & 1.02 $\pm$ 0.02\\
    LiCl        & 6.03 $\pm$0.80$\times$10$^{-3}$        & 83    & 0.682 $\pm$ 0.090\\
    CaCl$_{2}$  & 1.25$\pm$0.1$\times$10$^{-2}$    & 149   & 0.789 $\pm$ 0.063\\
    MgSO$_{4}$  & 8.80$\pm$1.2$\times$10$^{-3}$   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 2.40$\pm$0.28$\times$10$^{-2}$   & 395   & 0.575 $\pm$ 0.063\\
    K$_{2}$SO4$_{4}$    & 1.98$\pm$0.20$\times$10$^{-2}$  & 258   & 0.722 $\pm$ 0.073\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{img/SI-eta-all.pdf}
  \caption{Measured $\eta$ values as function of $V_{\mathrm{G}}$ for
    various salts used in this work. The concentration in HCR is 0.1
    mM in all cases.}
  \label{fig:eta-all}
\end{figure}

\section{Control experiment with copper tape}
\label{sec:copper}
We make sure that the applied voltage is effectively applied via the
graphene membrane and not through leakage directly coupling the
copper tape to the ionic solution. For this purpose, we use a device as in \Fig
1a, omitting the graphene and measuring the current through the
membrane with the Autolab electrochemical workstation using the
membrane as working electrode and a platinum wire as counter/reference
electrode 3.5 cm apart in 0.1 mM KCl. When no graphene is inserted in
the device, a current in the baseline range ($\sim$ 0.1 nA) is
observed, indicating no current passing from the membrane. However, if
graphene is inserted, a current in the range of $\sim$ 10 - 100 nA is
measured. If the copper tape is in contact with the solution when
omitting the Kapton tape, the current increases to $\sim$ 1 - 10
$\mathrm{\mu}$A, independent of the exposed area of the copper tape to
the solution.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-resistance-rejection.pdf}
  \caption{Simple circuit analog model for the transport through
    PG-PCTE membrane. \textbf{a}. Schematic of PG-PCTE membrane in
    cross-section view. \textbf{b}. Model for transport resistance in
    analogy to an electric circuit. This schematic depicts a cut
    through a PG-PCTE membrane showing a single PCTE pore and multiple
    graphene pores.}
  \label{fig:R-model}
\end{figure}
The ionic transport through the PG-PCTE membrane is the concerted
effect the transport resistance through PG and PCTE membranes,
respectively. As schematically shown in \Fig{} \ref{fig:R-model}, the
overall transport resistance of the system is composed of the series
resistance $R_{\mathrm{G}}$ and $R_{\mathrm{PCTE}}$, respectively. For
both PG and PCTE membranes, the resistance $R$ is defined as the ratio
between the concentration between HCR and LCR $\Delta c$, and the total diffusive current $I$ as:
\begin{equation}
\label{eq:1}
R_{i} = \Bigg|\frac{\Delta c}{I_{i}}\Bigg| =  \Bigg|\frac{\Delta c}{J_{i}A_{i}}\Bigg|
\end{equation}
where $i$ denotes the membrane type and $A_{i}$ is the area of the
pore. We further treat $R_{\mathrm{G}}$ as the effective resistance of
$\gamma$ parallel resistors with individual resistance
$R_{\mathrm{g}}$, where $\gamma$ is the average number of PG pores per
PCTE channel, and $R_{\mathrm{g}}$ is the average transport resistance
per PG pore. The value of $\gamma$ is determined from geometric
analysis of both membranes: given the pore number density of PG
($\sim$1.25$\times$10$^{14}$ m$^{-2}$), the pore number density of
PCTE ($\sim$1.50$\times$10$^{12}$ m$^{-2}$), and the percentage of
PCTE pores with PG on top ($\sim$0.13), $\gamma$ is determined to be
$\sim$10.83.

The objective of the circuit analog model is to find effective salt
rejection ratio through PG, $\xi$ extracted from the experimental
$\eta$ values (defined in main text Eq. \ref{main-eq:rejection}). By
defining $R_{\mathrm{G}}^{0}=R_{\mathrm{G}}(V_{\mathrm{G}}=0)$,
$\delta = R_{\mathrm{PCTE}} / R_{\mathrm{G}}^{0}$,
$\chi = R_{\mathrm{G}}/R_{\mathrm{G}}^{0}$, $\eta$ is rewritten using
transport resistances as:
\begin{equation}
  \begin{aligned}[t]
      \label{eq:eta-R}
  \eta &= 1 - {\displaystyle
    \frac{(R_{\mathrm{PCTE}} + R_{\mathrm{G}})^{-1}}
    {(R_{\mathrm{PCTE}} + R_{\mathrm{G}}^{0})}} \\
    &= 1 - \frac{\delta + 1}{\delta + R_{\mathrm{G}}/R_{\mathrm{G}}^{0}} \\
    &= 1 - \frac{\delta + 1}{\delta + \chi}
  \end{aligned}
\end{equation}
Consequentially, the
salt rejection ratio $\xi$, is associated with $\chi$ by
$\xi = 1 - \chi^{-1}$ and is related with $\delta$ and $\eta$ via:
\begin{equation}
  \label{eq:xi-deriv}
  \begin{aligned}[t]
    \xi &= 1 - \chi^{-1} \\
    &= \frac{(\delta + 1) \eta}{\delta\eta + 1}
  \end{aligned}
\end{equation}
which is main text Eq. \ref{main-eq:xi-def}. When $0<\eta<1$ (salt
rejection regime) and $\delta>0$, Eq. \ref{eq:xi-deriv} always
guarantees $\xi>\eta$, since $\delta + 1$ is always larger than
$\delta\eta + 1$. In other words, due to the existence of
$R_{\mathrm{PCTE}}$, the effective salt rejection through PG upon
gating is always larger than that experimentally observed.

Next we seek the value of $\delta$, the resistance ratio
between the PCTE and PG membranes, without gating. The range
of $\delta$ can be estimated from experimental data in main text
\Fig{} \ref{main-fig:2}b using the relation:
\begin{equation}
  \label{eq:exp-delta}
  \delta = \left(\frac{\boldsymbol{J}_{\mathrm{PCTE}}}{\boldsymbol{J}_{\mathrm{PG}}^{0}} -1 \right)^{-1}
\end{equation}
by assuming the validity of the circuit analog model for all salt
systems. As shown in \Fig{} \ref{fig:delta-compare}, the
experimentally estimated value of $\delta$ varies from 1 to 5,
indicating that the resistance of the PG membrane is dominating in the
PG-PCTE system.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.55\linewidth]{img/SI-delta-estimate.pdf}
  \caption{Comparison between the $\delta$ value estimated from simple
    geometric model (Eq. \ref{eq:delta-resistance}) and from
    experimental value (Eq. \ref{eq:exp-delta}), showing good
    agreement between the two methods.  }
  \label{fig:delta-compare}
\end{figure}
$\delta$ can also be estimated using geometric parameters of both
membranes. For a long PCTE channel with length $L$ and
$r_{\mathrm{PCTE}}$, $R_{\mathrm{PCTE}}$ has power law of
$L/r_{\mathrm{PCTE}}^{2}$, while for a graphene pore
with radius $r_{\mathrm{G}}$ and vanishing thickness, the resistance of individual pore $R_{\mathrm{g}}^{0}$
has power law $1/r_{\mathrm{G}}$ \cite{O_Hern_2014}, which are expressed as:
\begin{subequations}
\begin{eqnarray}
  \label{eq:R-both}
  R_{\mathrm{PCTE}} &= &k_{\mathrm{PCTE}} \frac{L}{r_{\mathrm{PCTE}}^{2}} \\
  R_{\mathrm{g}}^{0} &= &k_{\mathrm{g}} \frac{1}{r_{\mathrm{G}}}
\end{eqnarray}
\end{subequations}
where $k_{\mathrm{PCTE}}$ and $k_{\mathrm{g}}$ are the coefficient
associated with PCTE and PG membranes, respectively. The value of
$\delta$ is then expressed as:
\begin{equation}
  \label{eq:delta-resistance}
  \delta
  = \frac{R_{\mathrm{PCTE}}}{R_{\mathrm{G}}^{0}}
  = {\displaystyle
    \frac{k_{\mathrm{PCTE}}}{k_{\mathrm{g}}}
    \frac{L r_{\mathrm{G}} \gamma}
        {r_{\mathrm{PCTE}}^{2}}}
\end{equation}
Here we take $L$=24 $\mu$m, $r_{\mathrm{G}}$=10 nm,
$r_{\mathrm{PCTE}}$=200 nm. From simple Fick's law,
the coefficient $k_{\mathrm{PCTE}}$ is determined as
$1/(D_{\mathrm{\pm}} \pi)$. On the other hand $k_{\mathrm{G}}$ varies
by the choice of model, ranging from $3 \pi / D_{\mathrm{\pm}}$ using
the analog of the Sampson formula for
fluid\cite{Roscoe_1949,Celebi_2014}, to $1/(2 D_{\mathrm{\pm}})$ from
the Hill formula for disk absorption model
\cite{Hill_1975,Grebenkov_2018}. As a result the theoretical value of
$\delta$ has lower bound of $\sim$2.2 (Sampson formula, orange line in
\Fig{} \ref{fig:delta-compare}) to $\sim$40.5 (Hill formula). We note
our experimental values of $\delta$ is in good agreement estimated
from the Sampson formula (which we use in the main text) while lower
than the Hill formula. Having a higher estimated value of $\delta$
could yield unreasonably high value of $\xi$ and the influence of
concentration is deviated from our theoretical predictions. Further
theoretical analysis taking nonidealities including surface adsorption
and chemical nature of pore edge, is required to have better
understanding of graphene's transport resistance.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    $c_{0}$ (mM) & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm)\\
    \hline
    0.1&        1.90$\times$10$^{-2}$ &        30.4\\
    0.33&       5.41$\times$10$^{-2}$ & 16.0\\
    1   &1.33$\times$10$^{-1}$ & 10.0\\
    3.3&     5.13$\times$10$^{-1}$ &  6.0\\
    10&      1.08    &3.3\\
    \hline
  \end{tabular}
  \caption{Conductivity and calculated Debye lengths of various KCl
    concentrations in order to obtain rejection versus
    concentration curves. }
  \label{tab:debye}
\end{table}

@@

In this section we provide a more detailed discussion about the
different ionic pathways through the graphene nanopore upon gating in
main text \Fig{} \ref{main-fig:4}b. As discussed in the main text, the
ionic migration arises from both diffusion (caused by
concentration gradient) and electrostatic drift (caused by electric
field).  The effect of salt rejection can be viewed as follows: when
$V_{\mathrm{G}}$ = 0, the flux $\boldsymbol{J}_{\mathrm{G}}$ is solely
contributed by the concentration gradient, while after applying a
positive $V_{\mathrm{G}}$, distinct pathways for cations /
anions are created and the drift flux counteracts the diffusion.

Applying a positive $V_{\mathrm{G}}$ changes the surface distribution
of the cations / anions. To see this, we plot the concentrations of
cations ($c_{+}$) and anions ($c_{-}$) near a graphene nanopore with
$r_{\mathrm{G}}$ = \unit[10]{nm}, $c_{0}$ = \unit[0.1]{mM} and $V_{\mathrm{G}}$ =
\unit[0.75]{V}, as shown in \Fig{} \ref{fig:conc}a and \ref{fig:conc}b,
respectively. The cations are depleted from the graphene surface,
while the anions are accumulated due to positive
$\psi_{\mathrm{G}}$. The anion concentration at the graphene surface
is enhanced $\sim{}$35 times compared with the bulk concentration in
HCR. The thickness of the depletion / accumulation layer is affected
by the bulk concentration. Here we show this by comparing the radial
distribution of $c_{+}$ and $c_{-}$ along the $z = 0$ line within the
graphene pore. \Fig{} \ref{fig:conc-r} shows such radial ionic
concentration distribution at different $c_{0}$ and $V_{\mathrm{G}}$
levels. When $c_{0}$ = \unit[0.1]{mM}, $\lambda_{\mathrm{D}}$ is larger than
$r_{\mathrm{G}}$, and the external gate voltage $V_{\mathrm{G}}$ has
control over both $c_{-}$ and $c_{+}$ within the graphene nanopore
(\Fig{} \ref{fig:conc-r}a and \ref{fig:conc-r}c). On the other
hand, when $c_{0}$ = \unit[100]{mM}, $\lambda_{\mathrm{D}}$ is much smaller
than $r_{\mathrm{G}}$, and the concentration is only affected by
$V_{\mathrm{G}}$ close to the pore edge (\Fig{} \ref{fig:conc-r}b
and \ref{fig:conc-r}d). Therefore by increasing $c_{0}$ (and
equivalently reducing $\lambda_{\mathrm{D}}$), $V_{\mathrm{G}}$ loses
control over the transport in the center of the pore and leads to a
lower $\xi$ value. Due to the existence of graphene's quantum
capacitance, the surface potential $\psi_{\mathrm{G}}$ is essentially
smaller than $V_{\mathrm{G}}$, and the maximum surface anion
concentration ($c_{0}$ = \unit[100]{mM}, $V_{\mathrm{G}}$ = \unit[1.25]{V}) is
$\sim{}$ \unit[3]{M}, much smaller than the saturated surface adsorption
density (at the order of \unit[10$^{2}$]{M}
\cite{bard_electrochemical_1980}). Therefore we expect the surface
enhancement of the anions under the FEM conditions to be realistic and
the influence of the anion concentration on the diffusivity is minimal
\cite{Tang_1999_I}.

Next we analyze the diffusion and drift contributions to the ionic
transport. As shown in \Fig{} \ref{fig:potential}, the
electrochemical potential energy $\mu_{i}$ is decomposed into the
electric potential energy $z_{i} c_{i} \psi_{i}$ (\Fig{}
\ref{fig:potential}a and \ref{fig:potential}b) and the chemical
potential energy $k_{\mathrm{B}}T \ln x_{i}$ (\Fig{}
\ref{fig:potential}c and \ref{fig:potential}d) near a graphene
nanopore with $r_{\mathrm{G}}$ = \unit[10]{nm} and $V_{\mathrm{G}}$ = \unit[0.75]
{V}. The transport of cations is dominated by its chemical potential and
the diffusion is most pronounced in the center of the pore. The electric and diffusive potentials are similar in magnitude, but with
opposite signs, counteracting each other. As a result, different pathways for cations
and anions emerge, where cations preferably pass through the pore
center, while anions mainly pass near the pore edge, as shown in
\Fig{} \ref{fig:flux}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-flux.pdf}
  \caption{z-component fluxes of cation ($J_{z+}$, \textbf{a}.)  and
    anion ($J_{z-}$ \textbf{b}.) near the graphene nanopore
    corresponding to \Fig{} \ref{main-fig:4}b in main text. The
    pathways for the different ions can be spatially distinguished:
    the flux of cation passes mostly through the center of the
    graphene nanopore while the anion passes mainly through the
    edge. The analysis is in consistent with the gradient of chemical
    potential as shown in the main text \Fig{} \ref{main-fig:4}b.}
  \label{fig:flux}
\end{figure}
